%%%%%%%%%%%%%%% Page Setup %%%%%%%%%%%%%%%

\documentclass[a4paper, 11pt]{book}
\usepackage[utf8]{inputenc}
\usepackage[framemethod=tikz]{mdframed}
\usepackage[margin=3cm]{geometry}
\setlength\parindent{0pt}
\setlength{\parskip}{\baselineskip}%
\usepackage{graphicx}
\usepackage{tgadventor}
%\renewcommand*\familydefault{\sfdefault}
\usepackage[T1]{fontenc}

%%%%%%%%%%%%%%% Math Stuff %%%%%%%%%%%%%%%

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{tikz}
\usepackage{tikz-cd}
\usepackage{tkz-euclide}
\usepackage{pgf-pie}
\usepackage{cancel}

%%%%%%%%%%%%%%% Quotation Stuff %%%%%%%%%%%%%%%

\usepackage[english]{babel}
\usepackage[autostyle]{csquotes}

%%%%%%%%%%%%%%% Color %%%%%%%%%%%%%%%

\definecolor{p}{rgb}{0.9118, 0.4961, 0.6804}
\definecolor{p1}{rgb}{1, 0.92, 0.963}
\definecolor{g}{rgb}{0.023, 0.477, 0.219}
\definecolor{g1}{rgb}{0.03, 0.75, 0.20}
\definecolor{g2}{rgb}{0.95, 1, 0.95}

%%%%%%%%%%%%%%% Link Setup %%%%%%%%%%%%%%%

\usepackage{hyperref}
\hypersetup{colorlinks=true, linkcolor=g, filecolor=g, urlcolor=g,}

%%%%%%%%%%%%%%% Box %%%%%%%%%%%%%%%

\newmdenv[innerlinewidth=2pt, roundcorner=4pt,linecolor=g1,innerleftmargin=20pt,
innerrightmargin=20pt,innertopmargin=20pt,innerbottommargin=20pt,backgroundcolor = g2]{mybox}

%%%%%%%%%%%%%%% Border Stuff %%%%%%%%%%%%%%%

\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}
\fancyhead[LE,RO]{\leftmark}
\fancyhead[RE,LO]{\thepage}
\fancyfoot[CE,CO]{}
\fancyfoot[LE,RO]{}

%%%%%%%%%%%%%%% Bug Fix Adjustments %%%%%%%%%%%%%%%

\setlength{\headheight}{14pt}
\setlength{\footskip}{55pt}

%%%%%%%%%%%%%%% Index %%%%%%%%%%%%%%%

\usepackage{makeidx}
\makeindex

%%%%%%%%%%%%%%% Custom Commands %%%%%%%%%%%%%%%

\def\greenlozenge{\mathbin{\color{g}\blacklozenge}}

\newcommand{\emphasis}[1]{\underline{\textbf{#1}} }
\newcommand{\vocab}[1]{\underline{\textbf{#1}}\index{#1}}

\newcommand{\claim}{\textbf{\underline{Claim}} }
%\newcommand{\corollary}{\underline{\textbf{Corollary}} }
\newcommand{\defn}{\underline{\textbf{Def}} }
\newcommand{\digression}{\underline{\textbf{Digression}} }
\newcommand{\easy}{\underline{\textbf{Easy}} }
\newcommand{\example}{\underline{\textbf{Example}} }
\newcommand{\fact}{\underline{\textbf{Fact}} }
\newcommand{\facts}{\underline{\textbf{Facts}} }
\newcommand{\goal}{\underline{\textbf{Goal}} }
\newcommand{\idea}{\underline{\textbf{Idea}} }
%\newcommand{\lemma}{\underline{\textbf{Lemma}} }
\newcommand{\need}{\underline{\textbf{Need}} }
\newcommand{\note}{\underline{\textbf{Note}} }
\newcommand{\proof}{\underline{\textbf{Proof}} }
\newcommand{\recall}{\underline{\textbf{Recall}} }
\newcommand{\remark}{\underline{\textbf{Remark}} }
%\newcommand{\theorem}{\underline{\textbf{Theorem}} }
\newcommand{\verify}{\underline{\textbf{Verify}} }


%%%%%%%%%%%%%%% New Theorems %%%%%%%%%%%%%%%


\newtheorem{theorem}{Theorem}[section]
\newtheorem{corollary}{Corollary}[theorem]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{conjecture}[theorem]{Conjecture}

%%%%%%%%%%%%%%% Math Operators %%%%%%%%%%%%%%%
\DeclareMathOperator{\A}{\mathbb{A}}
\DeclareMathOperator{\Aut}{Aut}
\DeclareMathOperator{\C}{\mathbb{C}}
\DeclareMathOperator{\characteristic}{char}
\DeclareMathOperator{\cod}{cod}
\DeclareMathOperator{\dom}{dom}
\DeclareMathOperator{\Fix}{Fix}
\DeclareMathOperator{\Frac}{Frac}
\DeclareMathOperator{\Free}{Free}
\DeclareMathOperator{\id}{id}
\DeclareMathOperator{\Ima}{Im}
\DeclareMathOperator{\Mor}{Mor}
\DeclareMathOperator{\N}{\mathbb{N}}
\DeclareMathOperator{\Obj}{Obj}
\DeclareMathOperator{\Q}{\mathbb{Q}}
\DeclareMathOperator{\R}{\mathbb{R}}
\DeclareMathOperator{\Stab}{Stab}
\DeclareMathOperator{\Span}{span}
\DeclareMathOperator{\Spec}{Spec}
\DeclareMathOperator{\Z}{\mathbb{Z}}
%%%%%%%%%%%%%%% Title %%%%%%%%%%%%%%%

\title{Transcendental Numbers (Rough Outline 9/28/23)}
\author{Vincent Lin}
\date{Fall 2023}


\begin{document}
\frontmatter
\maketitle
\tableofcontents
\mainmatter{}
\chapter{Preliminaries}
\section{Algebraic and Transcendental Numbers}

\defn{An \vocab{algebraic number} is a complex number that is the root of a finite nonzero polynomial in one variable with rational coefficients. A \vocab{transcendental number} is a complex number that is not algebraic.}\par

\note{We will use $\A$ to denote the set of algebraic numbers and we will use $\C \setminus \A$ to denote the set transcendental numbers. We will use $\R \setminus \A$ to denote the set of real transcendental numbers.}\par

To find some algebraic numbers, we can take a nonzero polynomial with rational coefficents and find its roots. By definition, these roots are algebraic numbers. For example, $\sqrt{2}$ is algebraic because it is a root of $x^2 - 2$. Also, $i$ is algebraic because it is the root of $x^2 + 1$. All rational numbers are algebraic as well. Let $\frac{p}{q} \in \Q$ be rational, where $p, q \in \Z$ and $q$ is nonzero. Then, it is the root of $x - \frac{p}{q}$.\par

What about transcendental numbers? Do they exist?\par

\begin{mybox}
    \theorem{Yes, transcendental numbers exist.}
\end{mybox}

\proof{Consider the set of algebraic numbers, which we will denote by $\A$. This set is countable. We will show this by forming a surjection  \[\phi : \left(\N \times \bigcup\limits_{n \in \N} \Q^{n}\right) \to \A.\] Note that $\N \times \bigcup\limits_{n \in \N}$ is countable because it is the Cartesian product of a countable set with a countable union of countable sets. By the fundamental theorem of algebra, we know that a polynomial of degree $k$ has $k$ (not necessarily distinct) roots. Therefore, we can number the $k$ roots from $1$ to $k$ for each polynomial. Thus, if we specify the nonzero rational coefficients $(a_0, \ldots, a_k)$ and an index $i$ for $i \in [k]$ to be the $i$-th root of the polynomial $a_k x^k + \cdots + a_0$, we get an algebraic number. We define the map $\phi(i, (a_0, \ldots, a_k))$ to be the $i$-th root of $a_k x^k + \cdots + a_0$ if it exists. Otherwise, we return zero. From the definition, we know that every algebraic number can be encoded this way since it is one of the roots of a nonzero rational polynomial. Thus, for all $a \in \A$, there must be an input $x$ such that $\phi(x) = a$, making this a surjective map from a countable set to $\A$, making $\A$ countable. However, since $\C$ is uncountable, it cannot be the case that all complex numbers are algebraic. Thus, if a complex number is not algebraic, it must be transcendental. Furthermore, $\R$ is uncountable so there must be real transcendental numbers as well.\[ \greenlozenge \]}

If transcendental numbers exist, then can we find an example? To show that a number is algebraic, we just have to find a nonzero rational polynomial that it is a root of, evaluate that polynomial at that number, and verify that we get zero. However, checking if a number is transcendental is a very hard problem. The rest of this paper will discuss the transcendence of a large class of numbers and methods for determining the transcendence.\par

Before that, we will generalize the concept of transcendence to other fields. In order to do this, we will list several definitions. Let $k, l$ be fields and let $l/k$ be a field extension.\par

\begin{itemize}
    \item{$a \in l$ is \vocab{algebraic} over $k$ if there is a nonzero polynomial $f \in k[x]$ such that $f(a) = 0$. Otherwise, $a$ is \vocab{transcendental} over $k$. In other words, an algebraic number is a complex number that is algebraic over $\Q$ and a transcendental number is a complex number that is not algebraic over $\Q$.} 
    \item{$X \subseteq L$ is \vocab{algebraically independent} over $k$ if for all $a_1, \ldots, a_t$ distinct and all $f \in k[x_1, \ldots, x_t]$, $f(a_1, \ldots, a_t) = 0$ implies $f = 0$.}
    \item{If $a \in l$ is algebraic over $k$, the \vocab{minimal polynomial} of $a$ over $k$, denoted $m_a$, is the unique monic irreducible polynomial which generates the kernel of the evaluation map $\phi(f) = f(a)$ for $f \in k[x]$. The \vocab{degree} of $\alpha \in k$ is the degree of the minimial polynomial $m_a$.}
    \item{$l/k$ is an \vocab{algebraic extension} if every element in $l$ is algebraic over $k$.}
    \item{$l$ is \vocab{algebraically closed} if every nonconstant single variable polynomial in $l$ has a root in $l$.}
    \item{An \vocab{algebraic closure} of $k$ is an algebraic extension of $k$ that is algebraically closed.}
\end{itemize}



\section{Liouville Numbers}
Now we will start looking for our first transcendental number! First, we will prove a small lemma.\par

\begin{mybox}
    \lemma{For any polynomial $p(x)$ with a root at $x = \alpha$ of multiplicity $m$, $p'(x)$ has the root $\alpha$ with multiplicity $m-1$.}
\end{mybox}

\proof{From the fundamental theorem of algebra, we can write $p(x)$ as ${(x-\alpha)}^{m}q(x)$ where $x-\alpha \nmid q(x)$. Now, taking the derivatives of both sides and using the product rule, we get 
\begin{align*}
    p(x) &= {(x-\alpha)}^{m}q(x) \\
    p'(x) &= \left({(x-\alpha)}^{m}\right)'q(x) + {(x-\alpha)}^{m}q'(x).
\end{align*}

\[ \greenlozenge \]}


\defn{A \vocab{Liouville number} is a real number $x$ such that for all $n \in \N$, there exists integers $p, q$ with $q > 1$ such that \[0 < \left\vert x - \frac{p}{q} \right\vert < \frac{1}{q^n}.\]}\par

Essentially, Liouville numbers are real numbers that can be approximately really, really closely by a sequence of rationals of the form $\left\{\frac{p_i}{q_i}\right\}$, where $q_i > 1$, and the distance between $x$ and $\frac{p_i}{q_i}$ is nonzero, but less than $\frac{1}{{(q_i)}^{i}}$.\par

\begin{mybox}
    \theorem{(Liouville's Approximation Theorem) Let $\alpha \in \R \setminus \Q$ be an algebraic number of degree $n$. Then, for any rational approximation $\frac{p}{q}$ to $\alpha$, we have \[\left\vert \alpha - \frac{p}{q} \right\vert \geq \frac{1}{q^n}\]}
\end{mybox}
\proof{Let $f$ be the minimal polynomial of $\alpha$ over $\Q$. Note that.}

blah blah blah
\section{Transcendence of \texorpdfstring{$e$}{LaTeX} and \texorpdfstring{$\pi$}{LaTeX}} 
\begin{mybox}
    \lemma{Let $f(x)$ be a real polynomial with degree $m$. Let \[
        I(t) = \int_{0}^{t} e^{t-x}f(x) dx.\] Then, \[I(t) = e^t\sum\limits_{j=0}^{m}f^{(j)}(0) - \sum\limits_{j=0}^{m} f^{(j)}(t).\] where $f^{(j)}(t)$ is the $j$-th derivative of $f$ with respect to $x$ evaluated at $t$.}\par
\end{mybox}

\proof{Probably induction and integration by parts. I will fill this in later.}
\chapter{Lindemann-Weierstrass Theorem}
\begin{mybox}
    \theorem{(Lindemann-Weierstrass Theorem) Suppose \[\alpha_1, \ldots, \alpha_n\] are algebraic numbers that are linearly independent over $\Q$. Then, \[ e^{\alpha_1}, \ldots, e^{\alpha_n} \] are linearly independent over the algebraic numbers. In other words, the extension field $\Q(e^{\alpha_1}, \ldots, e^{\alpha_n})$ has transcendence degree $n$ over $\Q$.}
\end{mybox}
\proof{}

\chapter{Gelfond-Schneider Theorem}
\section{Useful Lemmas}
In this chapter, we will prove the Gelfond-Schneider Theorem. We will introduce four lemmas to help prove this.

\begin{mybox}
    \lemma{Let the functions \[a_1(t), \ldots, a_n(t)\] be nonzero real polynomials $(\in \R[t])$ of degree \[d_1, \ldots, d_n\] respectively. Furthermore, let \[w_1, \ldots, w_n\] be (pairwise) distinct real numbers. Then, counting multiplicities, the function \[f(t) = \sum\limits_{j=1}^{n} a_j(t){e}^{{w_j}t}\] has at most $n - 1 + \sum\limits_{i=1}^{n} d_i$ real roots.}
\end{mybox}

\proof{Let $k = n + \sum\limits_{i=1}^{n} d_i$. We will prove this statement using strong induction on $k$.}

\begin{mybox}
    \lemma{Let $f(z)$ be an analytic function in the disk $D \subseteq \C$. Here, we define $D = \{z : \vert z \vert < d\}$ for some positive real $d$. Suppose $f$ is continuous on the closure of $D$, $\overline{D} = \{z : \vert z \vert \leq d\}$. Furthermore, let $\vert f \vert_d = \max\limits_{z \in \overline{D}, \vert z \vert = d} f(z)$. Then, for every $z \in \overline{D}$, $\vert f(z) \vert \leq \vert f \vert_d$.}
\end{mybox}
\proof{Seems very difficult. Involves Maximum Modulus Principle, proof is omitted in most texts.}
\begin{mybox}
    \lemma{something from complex analysis}
\end{mybox}

\begin{mybox}
    \lemma{something with matrices}
\end{mybox}


\section{Main Theorem}
\begin{mybox}
    \theorem{(Gelfond-Schneider Theorem) Let $\alpha$ and $\beta$ be algebraic numbers such that $\alpha \notin \{0, 1\}$ and $\beta \in \R \setminus \Q$. Then, ${\alpha}^{\beta}$ is transcendental.}\par
\end{mybox}

\proof{We will incorporate four lemmas in this proof}
\chapter{Baker's Theorem}
\begin{mybox}
    \theorem{(Baker's Theorem) Let $\mathbb{L} = \left\{\lambda \in \C : e^{\lambda} \in \overline{\Q}\right\}$. Then, if \[\lambda_1, \ldots, \lambda_n \in \mathbb{L}\] are linearly independent over $\Q$, }
\end{mybox}
\chapter{Schanuel's Conjecture}
\begin{mybox}
    \conjecture{(Schanuel's Conjecture) Suppose we have $n$ complex numbers \[z_1, \ldots, z_n\] that are linearly independent over $\Q$. Then, $\Q(z_1, \ldots, z_n, e^{z_1}, \ldots, e^{z_n})$ has transcendence degree at least $n$ over $\Q$.} 
\end{mybox}

\backmatter{}
\printindex
\end{document}
